Stuff

### 2014-01-23 paper/Introduction.tex
making it cleaner

\section{Introduction}
\label{sec:Introduction}

>>>>

:qh\section{Introduction}
\label{sec:Introduction}

Shared memory is a common abstraction for distributed systems as it
provides an easily understood model of interaction. To achieve this
abstraction efficiently, many hardware systems implement a cache
coherence implementation.

Since the correctness of this abstraction is paramount in the correct
functioning of the system, significant effort is placed in making sure
that the cache coherence engine is correct. In practice, this is done
in two steps. First, a cache coherence protocol is constructed and
verified. This protocol generally directly represents many of
the key concurrent aspects of an implementation, \eg reads and
writes to the system are broken into local atomic micro-operations, but
ignores implementation details \eg buffer sizing and memory access
times. Second, once the protocol has been verified it is handed to
hardware experts who fill in the missing implementation details and
realize it at the Register Transfer Level (RTL) of gates, wires, and
registers. This involves further breaking down of local micro-operations to
make it realizable in hardware. This effort is not a one time translation, as
performance properties are not easily estimated; a significant amount of
performance testing is requires to select a good implementation.

From a methodological perspective, this two step process represent a
significant methodological failure. We spent a significant amount of
time verifying the protocol and then invalidate any correctness
guarantees to the implementation by performing not only a non-trivial
translation from the protocol representation to hardware, but also
the refinement to add missing implementation details.

Because verified protocols are often described via Term Rewrite Systems (TRS).
it was shown by Dave \etal{} that a verified protocol could be transliterated
into Bluespec, a hardware description language based on Guarded Atomic Actions,
a restriction of TRS \cite{Hoe:TCAD, DNA:CoherenceImplementation}. This reduces
the correctness of the implementation to the correctness of the protocol,
effectively removing any implementation bugs. Though in the current versions of
the Bluespec language, each atomic action corresponds to one hardware clock
cycle, there has been some work to reduce this restriction \cite{Karczmarek}.
%This requires that the protocol description includes many important
%implementation-level details which had previously been elided for simplicity
%of the description. Practically, beyond some additional descriptional noise in
%the proof this is not a show stopper.

The issue with this protocol transliteration approach is that design of a
hardware system requires many design tweaks to reach optimal performance as one
cannot practically know the correct low-level hardware results without
evaluating the low-level representation.
Each of these changes can potentially breaks
correctness and a full reverification of correctness is needed in
principle. As hardware designers are unlikely to have the expertise or
skill this effectively nullifies the verification guarantee in design.

To fully realize this approach of design we need to provide not just a
single proof of correctness, but a proof framework covering any
variation of a design that might be of interest for performance
reasons. This would enable implementation level changes without
compromising on the ability to verify correctness of the design.

In practice, protocol variants rely on a small set of general ``local building
block'' properties, \eg{} messages always get from source to destination or
that the local cache's reaction to receiving a protocol message results in the
appropriate messages being sent, the appropriate local state being changed \etc{}.
% about parts of the
%design to be true for the full system correctness.
These properties are easily verifiable since they correspond to a small
partition of the system, for instance a single cache.
The details of how these properties hold vary depending on the precise design
variant. For instance, the reason why some property holds, like the
point-to-point FIFO ordering for messages transferred between two nodes,
depends on the type of network; the proof will be very different for a mesh
network compared to a point-to-point network.

In this paper, we describe an approach to provide this proof framework for any
particular cache coherence protocol via the example of a MSI-style protocol and
show how it can be extended to other protocols like MESI and MOESI. We show how
the modular structure of the proof and design can be shared and how
modification of micro-archtectural aspects of design directly translates to
just reverification of our local building block properties. We use Coq for our
proof framework. We also show how an atomic action based MSI protocol
description can be directly transliterated into an Inductive definition in Coq,
and verify its local building block properties formally, thus completing the
picture.
